---
title: "DATA 605 : ASSIGNMENT 2"
author: "Ramnivas Singh"
date: "09/05/2021"
output:
  html_document:
    theme: default
    highlight: espresso
    toc: no
  pdf_document:
    toc: yes
    toc_depth: '5'
    extra_dependencies: animate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```
--------------------------------------------------------------------------------
# Problem set 1 
 1. Show that $A^TA \neq AA^T$ in general. (Proof and demonstration.)

## Solution

The most simple situation is when matrix A has n rows and m columns and m does not equal n, then the resulting matrix will have m(2) columns and n(2) rows; the opposite situation (the matrix is transposed) will result in a different number of columns and rows.

Due to the order of multiplication in that also applies when matrix A * matrix B does not equal matrix B * matrix A, matrix A transposed * matrix A does not equal matrix A * matrix A transposed. 

```{r}
# Create a 3 x 3 matrix.
matrix1 <- matrix(c(1,3,0,4,1,5,2,0,7), nrow = 3, byrow = TRUE)
matrix1
# using R get the transpose for matrix1 matrices.
matrix1T <- t(matrix1)
matrix1T
#calculate both  ATA and AAT.
ATA <- matrix1T %*% matrix1
ATA
AAT <- matrix1 %*% matrix1T
AAT
#It's clear that ATA???AAT for this "general" matrix.
# check programmatically AAT == ATA
AAT == ATA
```


2. For a special type of square matrix **A**, we get $A^T*A = A * A^T$. Under what conditions could this be true?  (Hint:  The Identity matrix **I** is an example of such a matrix).

## Solution 

Since the transpose will equal the original matrix, symmetric matrices will have the transpose x matrix = matrix x transpose, lets try to prove it using R code.
```{r}
mat <- matrix(c(1,0,3,0,2,5,3,5,3), nrow=3, ncol=3) 

# Print Matric
mat

# Transpose Matric
t(mat)

mat%*%t(mat)

t(mat)%*%mat
```


# Problem set 2

Matrix factorization is a very important problem.  There are supercomputers built just to  do  matrix  factorizations.   Every  second  you  are  on  an  airplane,  matrices  are  being factorized.  Radars that track  ights use a technique called Kalman Filtering.  At the heart of Kalman Filtering is a Matrix Factorization operation.  Kalman Filters are solving linear systems of equations when they track your flight using radars.

Write  an  R  function  to  factorize  a  square  matrix  A  into  LU  or  LDU,  whichever  you prefer. You don't have to worry about permuting rows of A and you can assume that A is less than  5x5,  if  you  need  to  hard-code  any  variables  in  your  code.

## Solution
Use the built function to see that that LDU = A.

$A = L * U$

For a 3x3 matrix:

```{r}
factorization <- function(A){
  e21 <- matrix(c(1,-A[2,1]/A[1,1],0,0,1,0,0,0,1), nrow=3)
  matrixY <- e21 %*% A
  
  e31 <- matrix(c(1,0,-matrixY[3,1]/matrixY[1,1],0,1,0,0,0,1),nrow=3) 
  matrixY <- e31 %*% matrixY
  e32 <- matrix(c(1,0,0,0,1,-matrixY[3,2]/matrixY[2,2],0,0,1),nrow=3)
  U <- e32 %*% matrixY
  L <- solve(e21) %*% solve(e31) %*% solve(e32)
  
  A == L %*% U
}
A <- matrix(c(1,2,3,1,1,1,2,0,1),nrow=3)
factorization(A)
```