---
title: 'Data 608 : Module1'
author: "Ramnivas Singh"
date: "02/14/2022"
output:
  html_document:
    theme: default
    highlight: default
    toc: no
    toc_depth: 4
    toc_float:
      collapsed: yes
  word_document:
    toc: no
    toc_depth: '4'
  pdf_document:
    toc: no
    toc_depth: '4'
subtitle: Principles of Data Visualization and Introduction to ggplot2
---


**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```

```{r}

summary(inc)

```

```{r  results = FALSE, message = FALSE, warning = FALSE}
# Insert your code here, create more chunks as necessary
# Add required libraries
library(DT)
library(tidyverse)
library(ggplot2)
```
Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

```{r}

# Insert your code here, create more chunks as necessary
# Adding a table to seach and sort data in tabular form.
datatable(inc, 
          options = list(pageLength = 10,searching = TRUE, filter=TRUE, 
                         pageLength = FALSE),rownames = FALSE)
# Next I like calculate standard deviation and IQR to understand data skewness
sd(inc$Revenue)
sd(inc$Growth_Rate)
sd(inc$Employees, na.rm = TRUE)

#Lets do IQR in case the data is skewed
IQR(inc$Revenue)
IQR(inc$Growth_Rate)
IQR(inc$Employees, na.rm = TRUE)

```


## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r}

# Answer Question 1 here
# Ask : Give distributions of categorical data: States or Industries.
# We will use a bar which seems to be an intuitive way to present the information. Also bar will provide an comparable view in plot

state_counts <- inc$State %>% table() %>% as.data.frame(stringsAsFactors=FALSE)
colnames(state_counts) <- c('State', 'Frequency')
ggplot(state_counts, aes(x=reorder(State, Frequency),y=Frequency, color=State)) +
  geom_bar(stat='identity', color = 'white', fill=rainbow(52)) +labs(title="Distribution of Companies by State", x="US States", y="Number of Companies")+
  coord_flip() +theme(axis.text.y = element_text(angle = 5))
```

## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r  results = FALSE, message = FALSE, warning = FALSE}
# Answer Question 2 here
# By plot in question 1, we can see NY is 3rd state with most of the companies. 
NY <- subset(inc, State=="NY")
NY <- NY[complete.cases(NY), ]

ggplot(NY %>% group_by(Industry) %>% 
         summarise(`Median Employees` = median(Employees))) + 
  geom_col(
    aes(x=reorder(Industry, -`Median Employees`), y = `Median Employees`), 
    fill = "grey", 
    width = 0.25) +
  geom_text(
    aes(x = Industry, y = `Median Employees`, label=round(`Median Employees`, digits = 0)), 
    vjust=-0.25, 
    size=3,
    color="gray") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        axis.text.y = element_blank(),
        axis.title=element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.margin = margin(1, 1, 5, 45)
        ) +
  labs(title = "Median Number of Employees per Company by Industry - NY")
```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r}

# Answer Question 3 here
inc %>%
  filter(complete.cases(.)) %>%
  group_by(Industry) %>%
  summarise(Employees_n = sum(Employees),
            Revenue_n = sum(Revenue)) %>%
  mutate(Revenue_Per_Employee = Revenue_n / Employees_n) %>%
  ggplot(., aes(x= Revenue_Per_Employee, y = reorder(Industry, Revenue_Per_Employee))) +
  geom_bar(stat = "identity", ) +
  labs(title = "Distribution of Revenue/ Employee by Industry", y = "Industry in NY", 
       x = "Revenue/ Employee in NY")
```


