---
output:
  html_document:
    theme: default
    highlight: espresso
    toc: no
  pdf_document:
    toc: yes
    toc_depth: '5'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```

#### Problem 2 Page 321

Let X be a continuous random variable with values uniformly distributed over the interval [0, 20].
(a) Find the mean and variance of X.
(b) $P(|X - 10| \ge 2)$, $P(|X - 10| \ge 5)$, $P(|X - 10| \ge 9)$ and $P(|X - 10| \ge 20)$ exactly. How do your answers compare with those of
Exercise 1? How good is Chebyshev's Inequality in this case?

#### Answer :

(a) Chebyshev's Inequality mean and variance are 

$E(X)={(a+b)}/2$

$Var(X)={(b - a)^2}/12$

So Mean $E(X)$ is :
```{r}
print((0+20)/2)
```

and Variance $Var(X)$ is : 
```{r}
print(((20-0)**2)/12)
```

(b) $P(|X - 10| \ge 2)$ =$1-P(|X - 10| \lt 2)$

=$1-P(-2 \lt (X -10) \lt 2)$

=$1-P(-2+10 \lt X \lt 2 +10)$

=$1-P(8 \lt X \lt 12)$

=$1-[P(X \lt 12) -  P(X \le 8)]$

=$1-[12/20 - 8/20]$

```{r}
print(1-((12/20) - (8/20)))
```

$P(|X - 10| \ge 5)$

=$1-P(|X - 10| \lt 5)$

=$1-P(-5 \lt (X-10) \lt 5)$

=$1-P(-5+10 \lt X \lt 5+10)$

=$1-P(5 \lt X \lt 15)$

=$1-[P(X \lt 15) -  P(X \le 5)]$

```{r}
print(1-((15/20) - (5/20)))
```

$P(|X - 10| \ge 9)$

=$1-P(|X - 10| \lt 9)$

=$1-P(-9 \lt (X-10) \lt 9)$

=$1-P(-9+10 \lt X \lt 9+10)$

=$1-P(1 \lt X \le 19)$

=$1-[P(X \lt 19) -  P(X \le 1)]$

```{r}
print(1-((19/20) - (1/20)))
```

$P(|X - 10| \ge 20)$

=$1-P(|X - 10| \lt 20)$

=$1-P(-20 \lt (X-10) \lt 20)$

=$1-P(-20+10 \lt X \lt 20+10)$

=$1-P(-10 \lt X \le 30)$

=$1-P(0 \lt X \le 20)$

=$1-1$

```{r}
print(1-1)
```

How do your answers compare with those of Exercise 1? How good is Chebyshev's Inequality in this case?

Answer :

From problem 1 : 

a. $\epsilon=2$, the upper bound is $1$ from this problem upper bound is $0.8$
b. $\epsilon=5$, the upper bound is $1$ from this problem upper bound is $0.5$
c. $\epsilon=9$, the upper bound is $\approx 0.4115$ from this problem upper bound is $0.1$
d. $\epsilon=20$, the upper bound is $\approx 0.0833$ from this problem upper bound is $0.0$

Chebyshev estimates are in general not very accurate. We see variance for $\epsilon=2, 5, 9, 20$
