---
title: "DATA 605 : Week 11 - Linear Regression Model"
author: "Ramnivas Singh"
date: "11/07/2021"
output:
  pdf_document:
    toc: no
    toc_depth: '5'
  html_document:
    theme: default
    highlight: espresso
    toc: no
---

Lets use the MPG dataset. https://archive.ics.uci.edu/ml/datasets/Auto+MPG. We need to tidy the data.

```{r}
auto <- read.table(url("https://archive.ics.uci.edu/ml/machine-learning-databases/auto-mpg/auto-mpg.data"), header =FALSE)
```


```{r}
names(auto) <- c("mpg", "cylinders", 
                 "displacement", 
                 " horsepower", 
                 "weight", 
                 "acceleration", 
                 "model year",
                 "origin", 
                 "car name")
```


```{r}
auto.df<-data.frame(auto)
auto.df$X.horsepower <- as.numeric(as.character(auto.df$X.horsepower))
auto.df<-na.omit(auto.df)
summary(auto.df)
```

Check distribution of response variable 

```{r}
mpg <- auto.df$mpg 
h<-hist(mpg, breaks=10, col="red", xlab="Miles Per Gallon", 
    main="Histogram with Normal Curve") 
xfit<-seq(min(mpg),max(mpg),length=50) 
yfit<-dnorm(xfit,mean=mean(mpg),sd=sd(mpg)) 
yfit <- yfit*diff(h$mids[1:2])*length(mpg) 
lines(xfit, yfit, col="blue", lwd=2)
```

Build model on variables that do not include levels

```{r}
auto.df2 <- subset(auto.df, select = c(mpg, cylinders, displacement, X.horsepower, weight,acceleration))

```

Build a model 

```{r}
mod <- lm(mpg ~ ., data=auto.df2)
summary(mod)
```

Model Validation 
Residuals
```{r}
hist(mod$residuals);
qqnorm(mod$residuals);
qqline(mod$residuals)
```

Constant Variance
```{r}
library(olsrr)
ols_test_breusch_pagan(mod);
plot(fitted(mod), residuals(mod), xlab="fitted", ylab="residuals")
abline(h=0)
```

Based on these residuals, a linear model is not a good fit. The residuals demonstrate the need for some transformation on the response variable. The constant variance check is also a strong indicator that a linear model is not the way to go. 
