---
title: 'Data 607 : Assignment 3 - Character manipulation & Data processing'
author: "Ramnivas Singh"
date: "02/21/2021"
output:   
  html_document:
    theme: default
    highlight: espresso
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Overview

#1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"\

#2 Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"
Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

The two exercises below are taken from R for Data Science, 14.3.5.1 in the on-line version:
\
#3 Describe, in words, what these expressions will match:

(.)\1\1
"(.)(.)\\2\\1"
(..)\1
"(.).\\1.\\1"
"(.)(.)(.).*\\3\\2\\1"
#4 Construct regular expressions to match words that:
\
- Start and end with the same character.\
- Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)\
- Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)


# Preparation Work

## Source dataset
This step is to source the data set in R program from github location 

```{r}
library (readr)
dataUrl="https://raw.githubusercontent.com/rnivas2028/MSDS/Data607/Assignment3/majors-list.csv"
majorDataSet <- read.csv(dataUrl, header = TRUE, sep = ",", stringsAsFactors = FALSE)
```

# Excercise 1

Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"\

```{r}
key='DATA|STATISTICS'
majorSubDataSet <- majorDataSet$Major[grep(key, majorDataSet$Major)]
print(majorSubDataSet)
```

--------------------------------------------------------------------------------

\clearpage


# Excercise 2

Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"
Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")


```{r}
dataSet <- data.frame(c("bell pepper", "bilberry", "blackberry","blood orange","blueberry","cantalope","chili pepper","cloudberry","elderberry","lime","lychee","mulberry","olive","salal berry"))
cat(paste(dataSet), collapse=",")
```

--------------------------------------------------------------------------------

\clearpage

# Excercise 3

Describe, in words, what these expressions will match:

(.)\1\1
"(.)(.)\\2\\1"
(..)\1
"(.).\\1.\\1"
"(.)(.)(.).*\\3\\2\\1"

Answers:
(.)\1\1
This regular expression is used to match a pattern in a strings with character repeats in it.

“(.)(.)\2\1”
This regular expression is used to match strings with a set of 4 characters with 2 characters attached to the same 2 characters in reverse order(e.g: otto)

(..)\1
This regular expression is used to match any strings that have a repeated pair of letters

“(.).\1.\1”
This regular expression is used to match any strings that has the same character repeat 3 times and they are all separated by one character. (ex: papaya)

"(.)(.)(.).*\3\2\1"
This regular expression is used to match any strings with 3 characters followed by zero or more characters followed by the original 3 characters in reverse order.. (ex:abccba)

--------------------------------------------------------------------------------

\clearpage

# Excercise 4
Construct regular expressions to match words that:
\
- Start and end with the same character.\
- Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)\
- Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r}
dataSet <-c("cell","apple","dog","ada","bob","sense","church","banana","pepperoni","red","green","England","eleven","ten","twelve","soso","oso","bandana", "Louisiana", "Missouri", "Mississippi", "Connecticut", "google", "conscience","dalda","short","Evon","ele","Tort")
#4.1
expression="^(.)((.*\\1$)|\\1?$)"
result <- str_subset(dataSet,expression )
result

#4.2
expression="([A-Za-z][A-Za-z]).*\\1"
result <- str_subset(dataSet,expression )
result

#4.3
expression="([A-Za-z]).*\\1.*\\1"
result <- str_subset(dataSet,expression )
result

```


--------------------------------------------------------------------------------

\clearpage